---
title: "Shapeslosion data cleaning"
author: "Yuan Wang"
date: "5/11/2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Introduction

#### Goal
* We want to clean the raw Shapessplosion data. The goal is to filter out all rows that contain gender information, and then add a gender column with 0 indicating female and 1 be male.
    
#### Method
* We first filter out a gender data using v1label1.
* Then, we mutate a column to the gender data where if the user was male we denote as 1, and if female we denote as 0.
* We use a for loop to select all groupID's whose sample size is less then 50 and greater than 5. Also we use only groups that have more than 1 male and 1 female observations. (Otherwise t.test would fail).
* We save each group to an individual .csv file.
* We then bind all .csv file together and obtain a clean gender data.

```{r, include=FALSE}
library(plyr)
library(dplyr)
library(rmarkdown)
library(rprojroot)
library(mosaic)
library(data.table)
library(xlsx)
```

## Import and trim data

We use only untimed data with non-zero time used recording.

```{r}
Original <- read.csv("original_raw_data.csv")
shape <- Original
shape_untime <- shape[shape$requestedTime==0,]
shape_untimed <- shape_untime[shape_untime$timeUsed!=0,]
```

We changed some numerical variables to factors since they have only a fixed number of levels. We also changed the unit of the time data from milliseconds to seconds.

```{r}
shape_untimed$numShapes <- as.factor(shape_untimed$numShapes)
shape_untimed$matchingScheme <- as.factor(shape_untimed$matchingScheme)
shape_untimed$requestedTime <- as.factor(shape_untimed$requestedTime)
shape_untimed$timeUsed <- as.numeric(shape_untimed$timeUsed)
shape_untimed$timerDisplay <- as.factor(shape_untimed$timerDisplay)
shape_untimed$numErrors <- as.numeric(shape_untimed$numErrors)

shape_untimed <- mutate(shape_untimed, TimeUsedSec = shape_untimed$timeUsed/1000)
```

We further trimed the data to include a varible column named Gender. 

```{r}
gender <- filter(shape_untimed, tolower(strtrim(shape_untimed$v1label,3))=="gen" |
                   tolower(strtrim(shape_untimed$v1label,3))=="sex" |
                   tolower(strtrim(shape_untimed$v1label,4))=="male" |
                   tolower(strtrim(shape_untimed$v1label,1))=="f")

case <- (tolower(strtrim(gender$v1value,1)) == "m") & (tolower(strtrim(gender$v2value,1)) =="f")
gender1 <- gender[(strtrim(gender$v1value,1) != "1") &
                    (gender$v1value != "2") &
                    (gender$v1label != "Gender\rOrder") &
                    (gender$v1label != "female\rorder") &
                    (gender$v1value != "0") &
                    (tolower(gender$v1value) != "morf") &
                    (gender$studentID != "mb") &
                    (gender$studentID != "31207") &
                    (gender$v1value != "cat") &
                    (gender$studentID != "3659") &
                    (tolower(gender$v1value) != "attempt") &
                    (!case),]

gender1 <- mutate(gender1, 
                  gender = as.factor(ifelse(
                    pmax((tolower(strtrim(gender1$v1label,1)) == "m"),
                         (tolower(strtrim(gender1$v1value,1)) == "m")), 
                    1, 
                    ifelse(pmax(tolower(strtrim(gender1$v1label,1)) == "f",
                                (tolower(strtrim(gender1$v1value,1)) =="f")), 
                           0, 
                           NA))))

gender2 <- na.omit(gender1)
```

## Extract GroupID 

Then, within the gender data, we identified the group names with sample size between 5 and 50. We make a vector of groupID of our desired groups.

```{r}
tb <- as.data.frame(table(gender2$groupID))

##Select groups that are under size 50
tb2 <- tb[tb$Freq >= 5 & tb$Freq <= 50,]

##Create a vector of groupID's whose size is between 25 and 50
selected_groupID <- as.character(tb2$Var1)
```


## Create .csv files

Using a for loop, we traverse through the vector of groupID and, for each group with the corresponding groupID, we write out the data to a .csv file.

```{r, echo=TRUE}
par(mar=c(1,1,1,1))
par(mfrow = c(5,5))
groupName <- c()
pvalues <- c()
for (i in 1:length(selected_groupID)) {
  female <- gender2[gender2$groupID == selected_groupID[i] & gender2$gender == 0,]$TimeUsedSec
  male <- gender2[gender2$groupID == selected_groupID[i] & gender2$gender == 1,]$TimeUsedSec
  if (length(female) > 1 & length(male) > 1) {
    write.csv(gender2[gender2$groupID == selected_groupID[i],], file =paste(selected_groupID[i], "group.csv"))
  }
}
```



## Create clean data

We then read each .csv file back into data files, and  then use rbind to bind them together to produce one clean data set.

```{r warning=FALSE}
data <- read.csv("hj190f14 group.csv")
data <- rbind(data, read.csv("MA376C1 group.csv"))
data <- rbind(data, read.csv("msp2013 group.csv"))
data <- rbind(data, read.csv("MA376I3 group.csv"))
data <- rbind(data, read.csv("MAT336S15 group.csv"))              
data <- rbind(data, read.csv("MA376B1 group.csv"))
data <- rbind(data, read.csv("MATH22015 group.csv"))
data <- rbind(data, read.csv("MATH22018 group.csv"))
data <- rbind(data, read.csv("mth22602 group.csv"))
data <- rbind(data, read.csv("stats2 group.csv"))
data <- rbind(data, read.csv("mth32601 group.csv"))
data <- rbind(data, read.csv("mth32602 group.csv"))
data <- rbind(data, read.csv("MAED550 group.csv"))
data <- rbind(data, read.csv("336S14 group.csv"))
data <- rbind(data, read.csv("MAT336S14 group.csv"))
data <- rbind(data, read.csv("mth22601 group.csv"))
data <- rbind(data, read.csv("psych3700 group.csv"))
data <- rbind(data, read.csv("USCOTS15 group.csv"))

write.csv(data, file = "cleaned_gender.csv")

```

